### 2022-11-16 definitions.tex
details

\begin{definition}
A {\em (resp. strongly) well-structured transition systems} (abbreviated as WSTS resp. SWSTS) 
is a TS
$(S, \rightarrow, \leq)$
equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the wqo is (resp. strongly) compatible with the transition relation, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq t_2$  and $s_1 \rightarrow s_2$ , there exists 
	$t_2 \in S$ with 
	$s_2 \leq t_2$ and $t_1 \rightarrow^{*} t_2$ 			
				(resp. $t_1 \rightarrow^{1} t_2$ ).
\end{definition}

@@

\begin{proposition}
If $S$ is an SWSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
\end{proposition}

\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.
\fi

@@

(ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
\end{fact}

@@

\begin{lemma}
% Lemma 3 ([1]) 
% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
 Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of a strongly well-structured transition
system, we can decide whether we can reach $I$ from $s$.
\end{lemma}

\begin{proof}
We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
Then the
decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
We have
$$I^{k+1} = I \cup pre(I^k ) = I \cup
\bigcup_{s' \in B}
pre(\uparrow \{s' \}).$$

Since $pre(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
Fact 3, we can compute a basis of $I^{k+1}$.
\end{proof}

@@

\subsubsection{TS resilience}

@@

\problemx{TS $k$-resilience problem}
{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
{$\forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

\problemx{TS bounded resilience problem}
{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

\subsubsection{WSTS resilience}

@@

\problemx{WSTS $k$-resilience problem}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

Analogously, we formulate the bounded resilience problem for WSTSs.

\problemx{WSTS bounded resilience problem}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

>>>>

\begin{definition}
A {\em (resp. strongly) well-structured transition systems} (abbreviated as WSTS resp. S-WSTS) 
is a TS
$(S, \rightarrow, \leq)$
equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the wqo is (resp. strongly) compatible with the transition relation, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq t_1$  and $s_1 \rightarrow s_2$ , there exists 
	$t_2 \in S$ with 
	$s_2 \leq t_2$ and $t_1 \rightarrow^{*} t_2$ 			
				(resp. $t_1 \rightarrow^{1} t_2$ ).
\end{definition}

@@

\begin{proposition}
If $S$ is an SWSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
\end{proposition}
\alain{étrange ces deux propositions}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.
\fi

@@

(ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
\alain{A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
\end{fact}

@@

\begin{lemma}
% Lemma 3 ([1]) 
% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
 Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of an effective strongly well-structured transition
system, we can decide whether we can reach $I$ from $s$.
\end{lemma}

\begin{proof}
We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
Then the
decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
We have
$$I^{k+1} = I \cup pre(I^k ) = I \cup
\bigcup_{s' \in B}
pre(\uparrow \{s' \}).$$
\alain{n'est-ce pas $I^{k+1} = I^k \cup pre(I^k ) = I^k \cup
\bigcup_{s' \in B}
pre(\uparrow \{s' \}).$ ?}
Since $pre(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
Fact 3, we can compute a basis of $I^{k+1}$.
\end{proof}

@@

\subsubsection{Resilience}

@@

\problemx{$k$-resilience problem}
{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
{$\forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

\problemx{Resilience problem}
{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

\subsubsection{Resilience}

@@

\problemx{$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

Analogously, we formulate the resilience problem for WSTSs.

\problemx{Resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
{$\exists k \geq 0 ~ \forall s' \in D ~ (s \rightarrow^* s') \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}

### 2022-11-16 proof.tex
details

\begin{theorem}
{\sc SWSTS $k$-resilience} is decidable.
\end{theorem}

@@

Formally, let us show that $\uparrow post(\uparrow post^*(s)) \subseteq \uparrow post^*(s)$.
Let $b \in \uparrow post(\uparrow post^*(s))$ 
there exists $a', a, b$ such that
$s \rightarrow^* a'$,
$a' \leq a$,
$a \rightarrow b'$,
and
$b' \leq b$.
% By monotonicity, there exists b'' ≥ b' such that a → b'' . Therefore, x → ∗ b'' and b' ≥ b, hence b ∈ ↓ Post ∗ (x).
\textcolor{red}{By downward compatibility 
there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.}
\textcolor{green}{où est cette hypothèse downward compatibility  ?}
\item Procedure 2 computes
 $post^{\leq n}(s) \cap J$
 until it finds an element not in $ I^k$.

@@

\begin{theorem}
{\sc SWSTS resilience} is decidable.
\end{theorem}

@@

Iteratively check whether $k$-resilience holds. 
If this is the case, return $k_{min} = k$. Otherwise check whether 
$I^{k+1} = I^k$. If so, return $-1$ (false), otherwise
continue.
The stop condition is decidable
and by Fact 4 also sufficient. 
\end{proof}

>>>>

\begin{theorem}
{\sc $k$-resilience} is decidable for effective WSTS with strong monotony.
\end{theorem}

@@

Formally, let us show that $\uparrow post(\uparrow post^*(s)) \subseteq \uparrow post^*(s)$.
Let $b \in \uparrow post(\uparrow post^*(s))$ 
there exists $a', a, b$ such that
$s \rightarrow^* a'$,
$a' \leq a$,
$a \rightarrow b'$,
and
$b' \leq b$.
% By monotonicity, there exists b'' ≥ b' such that a → b'' . Therefore, x → ∗ b'' and b' ≥ b, hence b ∈ ↓ Post ∗ (x).
\textcolor{red}{By downward compatibility 
there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.}
\alain{où est cette hypothèse downward compatibility  ?}
\item Procedure 2 computes
 $post^{\leq n}(s) \cap J$
 until it finds an element not in $ I^k$.

@@

\begin{theorem}
{\sc $k$-resilience} is undecidable for effective WSTS ???
\end{theorem}

\begin{theorem}
{\sc resilience} is decidable for WSTS with strong monotony.
\end{theorem}

@@

Iteratively check whether $k$-resilience holds. 
If this is the case, return $k_{min} = k$. Otherwise check whether 
$I^{k+1} = I^k$. If so, return $-1$ (false), otherwise
continue.
The stop condition is decidable
and by Fact 4 also sufficient. 
\alain{Fact 4 don't exist}
\end{proof}
